### 2019-11-06 Notes/Notes.tex
Added preliminary notes on how the python module will work later on

\documentclass[]{article}

\usepackage{amsmath}
\usepackage{bbm}

\newcommand{\argmax}[1]{\text{arg}\hspace{-2pt}\max\limits_{#1}}
\newcommand{\bvec}[1]{\boldsymbol{#1}}

\author{}

%opening
\title{Notes}

\begin{document}

\maketitle

\section{Main Idea}

Integer linear programming is pretty useful apparently. The problem boils down to solving a system of linear equations and inequalities. Just Google linear integer programming, and Wikipedia has a good example for the general type of problem that is being solved. The underlying problem is classically NP-hard, so there is an possibility to exponentially speed-up this entire set of problems with quantum annealing. The main idea here is to encode inequalities into a quantum annealer.

Quick overview, a quantum annealers solves a QUBO problem (which is basically Ising Model). The problem can be boiled down to the form $x Q x$ where $x$ is a vector of qubits (so they can only take values of 0 or 1 or some super-position). We have to ``program'' $Q$ to solve our problem.

Following https://arxiv.org/abs/1812.06917 we know 1) how to encode fixed-precision numbers into a set of qubits. 2) We also know how to encode higher order problems (\textit{e.g.} $J x^4 + Q x^2$). 3)  We also learned that in general, since QA can only solve for the global minimum, we must optimize the squared-residual of a set of equations / inequalities. Anyways, read the paper if you haven't already.

To encode an inequality of the form $A < y < B$ we can introduce an auxiliary qubit $y_a$ such that the logical qubit is $Y = (y, y_a)$ (where $y$ and $y_a$ are logical qubits representing fixed-precision numbers encoded by $n$ qubits). The inequality is then encoded as minimizing the following $C_1 Y^4 - C_2 Y^2 + C_3$ such that we get a mexican-hat potential. Basically this was inspired from spontaneous symmetry breaking. I'm sure this can be connected to solving scalar phi-4 theory if we reallllly want to go there. But tuning the $C$s will tune the radius of the degeneracy and therefore change the range of $A$ and $B$.  Also $C_3$ shifts the minimum to zero, which is useful if we want to keep the energy positive, and interpret it as the squared-residual. Travis suggested that this might be connected to ``slack variables'' (see https://arxiv.org/abs/1811.11538 that I haven't read yet).

I'm worried that $y_a$ has nonlinear dependence with $y$ as we walk along the degeneracy though, which in practice may be an issue. Though increasing the number of qubits used to represent a fixed-precision decimal solves this problem exponentially quickly.

\section{Implementation}

So I have been following up on @CCChen's notes and added a bit software to it.

It's is basically the same thing minor two differences.

\subsection{Integer problem optimization}
What we eventually want to solve is an inequality of the form
\begin{equation}\label{max-x}
	\vec x_0 = \argmax{\vec x} \left( \vec c \cdot \vec x\right) \, ,
\end{equation}
for
\begin{align}\label{condition-x}
	A \vec x - \vec b &\leq \vec 0 \, , &
	A &\in \text{Mat}(\mathbbm R, (N_x, N_e)) \, , &
	\vec b &\in \mathbbm{R}^{N_e} \, &
	\vec c &\in \mathbbm{R}^{N_x} \, , &
	\vec x &\in \mathbbm{N}^{N_x}\, ,
\end{align}
where the additional dimension $N_e$ is the number of inequalities we have.
Note that the inequality in the above equation is almost completely sufficient to characterize all problems because, for example a $\geq$ in a certain row can be obtained by multiplying the row with minus one and a $<$ can be obtained by an infinitesimal shift in $\vec b$.
Equalities are different, but, when we introduce the slack variable, we can also take care this.
For example, we know that for
\begin{equation}\label{slack-optimization}
	p >  \max\limits_{\vec x}\left(\vec c \cdot \vec x\right) \, , \qquad
	\vec x_0 = \argmax{\vec x} \left( \vec c \cdot \vec x - p f(\vec x) \right)
\end{equation}
for constrained $\vec x$ and
\begin{align}
	f(\vec x) &= \min\limits_{\vec s}\left[
		\left(\tilde A \vec x - \vec{\tilde{b}} + D \vec s\right) \cdot \left(\tilde A \vec x - \vec{\tilde{b}} + D \vec s\right)
	\right]
	\, , \\
	\vec s &\in \mathbbm{N}^{N_e} \, , \qquad
	D \in \text{Mat}(\{0, 1\}, (N_e, N_e)) \, .
\end{align}
More specifically $D$ is diagonal and it's entries are either one or zero (they are zero when we have an equality instead of an inequality).
The entries 
\begin{align}
	\tilde A = (10^{m} A) & \in \text{Mat}(\mathbbm Z, (N_x, N_e)) \, , &
	\vec{\tilde{b}} &= (10^m \vec{b}) \in \mathbbm{Z}^{N_e}
\end{align}
are rescaled with a factor such that they are integer valued.
This assumes that $A$ and $\vec b$ are given for finite precision.
Because they are integer valued, one can find a vector $\vec s$, such that $\forall \vec x \, \exists \vec s \, : \, \tilde A \vec x - \vec{\tilde{b}} + D \vec s = 0$.
Thus $f(\vec x) \geq 0$ and furthermore
\begin{equation}
	f(\vec x_s) = 0 \iff A \vec x - \vec{b} \leq 0 \, .
\end{equation}
From left to right is true because each component $s_i \geq 0$ thus $\tilde A \vec x - \vec{\tilde{b}}$ must be $\leq \vec 0$ and this is invariant under scaling with a positive number.
From right to left is true because if $A \vec x - \vec b$, this is true for their integer version as well.
And since $- \vec m = \tilde A \vec x - \vec{\tilde{b}} \leq \vec 0$ for a vector $\vec m \in \mathbbm{N}^{N_e}$, we can pick $\vec m = \vec s$.

Therefore, if $p$ is sufficiently large we know that the \eqref{slack-optimization} is maximal if $f(\vec x) = 0$ and the maximum of \eqref{slack-optimization} is thus equal to $\max\limits_{\vec x}(\vec c \cdot \vec x)$ under the specified constraints.

\subsection{Map to bits}
Next we want to code up $\vec x$ and $\vec s$ as bit vectors, e.g.,
\begin{equation}
	q = \sum_{n=0}^{N_b-1} 2^n \psi_n^{(q)} \, , \qquad \psi_n^{(q)} \in \{ 0, 1\}\, .
\end{equation}
Note that we in principle could have different $N_b$ for $x$ and $s$---I will discuss the reasoning behind that later.

With this definition, we can introduce a new basis and map between integers and bits such that
\begin{align}
	q &= \vec Q \cdot \bvec{\psi}^{(q)} \, , &
	\bvec{\psi}^{(q)} &\in \{ 0, 1\}^{N_b} \,, &
	\vec Q^T &= \left(1, 2, 2^2, \cdots,  2^{N_b -1} \right) \, .
\end{align}

With these definition we have that
\begin{align}
	\vec \xi 
	&= 
	\begin{pmatrix}
		\vec x \\
		\vec s 
	\end{pmatrix}
	= Q \bvec \psi^{\left( \vec \xi \right)} \, , &
	Q 
	&=
	\begin{pmatrix}
		\vec Q^T & \vec 0^T & \vec 0^T & \cdots & \vec 0^T \\
		\vec 0^T & \vec Q^T & \vec 0^T & \cdots & \vec 0^T \\
		\vdots   & \ddots   & \ddots   & \ddots & \vdots \\
		\vec 0^T & \vec 0^T & \vec 0^T & \cdots & \vec Q^T
	\end{pmatrix}\, .
	\, , & 
	\psi^{\left( \vec \xi \right)} & \in \left(\{0, 1\}^{N_b}\right)^{N_e + N_x} \, ,
\end{align}
in other words $Q$ is a $(N_e + N_x) \times N_b(N_e + N_x)$ matrix.

\subsection{Integer problem optimization using bits}
Now we can map \eqref{slack-optimization} to the new basis
\begin{align}
	\vec x_0 
	&= \argmax{\vec x}
	\left( \bvec \gamma \cdot \bvec \psi^{\left( \vec \xi \right)} - p \phi\left(\bvec \psi^{\left( \vec \xi \right)}\right)\right) \, , &
	\bvec \gamma &= \begin{pmatrix}
		\vec c \\ \vec 0
	\end{pmatrix} \cdot Q \in \mathbbm{R}^{(N_x + N_e )N_b} \, , 
\end{align}
which is in principle independent of the $\vec s$ pars and
\begin{align}\label{bit-penalty}
	\phi\left(\bvec \psi^{\left( \vec \xi \right)}\right)
	&=
	\min\limits_{\vec s}\left[
	\left(\tilde \alpha Q \bvec \psi^{\left( \vec \xi \right)} - \vec{\tilde{\beta}}\right) 
		\cdot \left(\tilde \alpha Q \bvec \psi^{\left( \vec \xi \right)} - \vec{\tilde{\beta}}\right)\right]
	\, , &
	 \vec{\tilde{\beta}} &= \begin{pmatrix}\vec{\tilde{b}}\\\vec 0 \end{pmatrix}
	\,,&
	\tilde \alpha &= \begin{pmatrix}
		\tilde A & 0 \\ 0  & D
	\end{pmatrix} \, , &
\end{align}
where the above quantities are split up in $\vec x$ and $\vec s$ space.

In the following context we use that the components of the bit vector are either one or zero and thus $\psi^{\left( \vec \xi \right)}_i = \psi^{\left( \vec \xi \right)}_i \psi^{\left( \vec \xi \right)}_i$.
The argument of equation \eqref{bit-penalty} simplifies to
\begin{equation}
	\psi^{\left( \vec \xi \right)}_\nu
	\left[
		(\tilde \alpha Q)_{i\nu} (\tilde \alpha Q)_{i\mu}
		-  \delta_{\nu\mu}\tilde{\beta}_i (\tilde \alpha Q)_{i\mu}
		- \delta_{\nu\mu} (\tilde \alpha Q)_{i\nu} \tilde{\beta}_i
	\right]
	\psi^{\left( \vec \xi \right)}_\mu
	+ \tilde{\beta}_i \tilde{\beta}_i \, .
\end{equation}
Because the we compute the location of the optimal value, we can neglect the constant shift $\tilde{b}_i \tilde{b}_i$.
Furthermore, we can combine the search for optimal $\vec x$ and $\vec s$ by a search for optimal $\vec \xi$ and hence optional $\bvec \psi^{\left( \vec \xi \right)}$.
This results in
\begin{align}
	\bvec \psi^{\left( \vec \xi \right)}_{0} 
	&=
	\argmax{\psi^{\left( \vec \xi \right)}}\left( 
		\bvec \psi^{\left( \vec \xi \right)} \cdot \Omega \bvec \psi^{\left( \vec \xi \right)}
	\right)
	\, , &
	\Omega_{\nu\mu}
	&=
	\delta_{\nu\mu} \gamma_i Q_{i\nu}
	- \tilde \alpha_{ij} Q_{j\nu} \tilde \alpha_{ik} Q_{k\mu}
	+ 2 \delta_{\nu\mu}\tilde{\beta}_i \tilde \alpha_{ij} Q_{j\mu}
\end{align}

\end{document}